Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2012-11-16 1.tex
Moving stuff around, mostly.

Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature. At the risk of explaining what might be no more than an an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms. \citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms. An example from German is shown in (\ref{fd}) below.

>>>>

Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature. At the risk of explaining what might be no more than an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms. \citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms. An example from German is shown in (\ref{fd}) below.

### 2012-11-16 2.tex
Moving stuff around, mostly.

Scientific observations do not necessarily arrive at the appropriate granularity for analysis.
Since the granularity of measurement and analysis is determined by prior hypotheses about the nature of the system under study, disputes are inevitable as scientific paradigms evolve. 
This study argues that there are both theoretical and empirical arguments to doubt the implicit hypothesis linking scalar wordlikeness judgements and a gradient model of phonotactics. First, intermediate judgements are characteristic of all gradient rating tasks, and therefore these judgements are irrelevant to the question of whether well-formedness is categorical or gradient.
Secondly, simple baselines better account for gradient well-formedness judgements than current computational models of phonotactic knowledge, suggesting that the gradience observed in these tasks do not derive from grammatical mechanisms.

@@

\section{Aspects in the theory of gradient grammaticality}

@@

\begin{figure}
\centering
\includegraphics{agg.pdf}
\caption{Subjects freely use intermediate ratings when asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively \citep{Armstrong1983}}
\label{agg}
\end{figure}

@@

Each step of this procedure is dubious, however.
First, speakers have difficulty perceiving \citep{Brown1956} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
Secondly, the computation of a scalar value serves no further purpose than to provide for gradient well-formedness judgements; even if this is the result of probabilistic parsing \citep[e.g.,][]{Coleman1997}, 
it need not be computed overtly.\footnote{
    An objection might be made here on evolutionary grounds. 
    It is still quite mysterious why the human language endowment includes constraints on pronominal binding, for instance, but these constraints are implicated in everyday language use.
    Far more bizarre is the suggestion that the linguistic endowment includes mechanisms only used in certain experimental tasks.}
Next, speakers must be able to consciously access and report the magnitude of this value (it must be \emph{cognitively penetrable} in the sense of \citealt{Pylyshyn1984}), an ability which is limited in many other domains.
Finally, \citet{Sprouse2011} argues that speakers do not (or cannot) scale well-formedness values to satisfy the assumptions magnitude estimation, a type of gradient rating task.
%\citep{Bard1996}

It is informative to compare this baroque architecture to a model of what is necessary to make a binary well-formedness judgement. When presented with a linguistic item in a judgement task, the grammar attempts to assign a parse. Speakers then access whether or not parsing was successful. 
There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}. 
As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures may not stored in linguistic memory. 
%This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
The fact that requests for repetition and clarification are ubiquituous in spontaneous speech illustrates further that speakers are frequently aware when parsing has failed.
Consequently, a great deal of evidence is needed to reject this simple model in favor of the gradient grammaticality architecture.

@@

It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared. 
\citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables. 
Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words.
Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, many of which are exceptionless, into probabilities. 
While this is consistent with their claim, that ``the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.

@@

\citet{Scholes1966} conducts several wordlikeness tasks with 7th-grade students (approximately 12--13 years of age). The data used here is his experiment 5, in which 33 speakers provide a ``yes'' or ``no'' as to whether each of the 63 items, presented aurally, are ``likely to be usable as a word of English''. Like the study by \citet{Albright2007}, the focus is on onset well-formedness and there is minimal diversity in rime type. Two items, [klʌŋ] \emph{clung} and [bɹʌŋ] \emph{brung} (a dialectical past participle of \emph{bring}), are excluded here as actual words of English. \citet{Albright2009a} and \citet{Hayes2008a} also use this data for the purposes of model evaluation; following \citet{Frisch2000}, they use the proportion of ``yes'' responses for each item so as to derive a continuous measure of well-formedness.

@@

Models are evaluated by comparing their scores to the average rating of each word using four correlation statistics. Each of these range between $[-1, 1]$, where $1$ indicates a perfect positive correlation and $-1$ denotes a perfect negative correlation. \citet{Hayes2008a} evaluate their model using the Pearson (``product-moment'') $r$, a parametric correlation measure.  \citet{Stevens1946} argues that statistics of this type are inappropriate for analysis of Likert scale data, like those used by \citet{Albright2007} and \citet{Albright2003b}.
The reason is that the Pearson $r$ makes a \emph{linearity assumption}. That is, it assumes that nonce words rated ``1'' and  ``3'', for instance, are just as different as those ``4'' and a ``6''. A weaker assumption, more appropriate for Likert scale data, is the \emph{monotonicity assumption}: that ``1'' is less English-like than ``3'', which is less English-like than ``4'', and so on. However, it also has been claimed that $r$ is particularly robust to violations of the linearity assumption \citep[e.g.,][]{Havlicek1976}. Pearson $r$ is reported here, but no stance is taken on its appropriateness for this data.
%\citet{Jamieson2004}
\citeauthor{Hayes2008a} also report Spearman $\rho$; this statistic requires only the weaker assumption of monotonicity, but it is difficult to give a simple interpretation to the coefficient.

Much easier to interpret are two other non-parametric statistics, the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ \citep{Noether1981}, as follows. These statistics are computed by comparing every model score/wordlikeness rating pair to every other: a comparison is counted as \emph{concordant} if the greater of the two model scores is the one associated with the greater of the two wordlikeness ratings (that is, the model ranks these two nonce words the in accordance with speakers' ratings), and as \emph{discordant} otherwise. These two statistics differ only in the treatment of ``ties'', pairs where either the model score or wordlikeness rating are identical. For $\gamma$, ties are ignored, and the coefficient is

@@

\noindent
where $c$ and $d$ represent the number of concordant and discordant pairs, respectively. The $\tau_b$ statistic uses a similar formula, but also incorporates a penalty for ties in model score which are not also paired with ties in wordlikeness ratings, or vis versa. \citet{Albright2009a} uses a variant of this statistic to evaluate wordlikeness models.

@@

Operationalizing ``phonotactic violation'' is somewhat more difficult. The simplest possible mechanism is chosen here: an onset or rime is identified as well-formed if it occurs with non-zero frequency in a representative sample, and is identified as ill-formed otherwise. The sample is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms, the latter though to be particularly strongly correlated with behavioral measures \citep{Brysbaert2009}. These pronunciations are then syllabified, and individual syllables parsed into onset and rime, according to a process described in detail in Appendix \ref{syllabification}. This is not is not to imply that all unattested onsets or rimes should be regarded as ill-formed, or that all onsets or rimes with non-zero frequency in this data are well-formed. For instance, \citet{Albright2009a} judges [dɹɛsp] \emph{dresp} to be phonotactically well-formed, despite the total lack of [ɛsp] rimes in English. Similar observations have been made concerning English onsets \citep[e.g.,][]{Cairns1972,Moreton2002}. In the other direction, there are precedents for labeling certain attested words as phonotactically ``peripheral'' (e.g., \citealt{Myers1987}, \citealt{Borowsky1989}), as lexical exceptions to language-specific prosodic constraints, without labeling these words as absolutely ungrammatical. Neither of these dissociations between attestation and well-formedness are implemented here, however.

@@

A second baseline is provided by measures of similarity to existing English words items, which has long been applied to model wordlikeness judgements (e.g., \citealt{Bailey2001}, \citealt{Greenberg1964}, \citealt{Kirby2007}, \citealt{Ohala1986b}, \citealt{Shademan2006,Shademan2007}, \citealt{Vitevitch1998,Vitevitch1999a}). \citeauthor{LSLT} (\citeyear{LSLT}: 151, fn.~27) suggests that grammaticality judgements in general might be influenced by similarity to existing grammatical structures, and \citet[417f.]{SPE} outline a similarity-based wordlikeness model. More recently, it has been observed (e.g., \citealt[51]{Coleman1997}, \citealt{Hay2004a}) that nonce words which flagrantly violate English sonority restrictions but which bear common affixes (e.g., *\emph{mrupation}) are rated highly English-like.

A wide variety of lexical similarity measures were considered, including a variant of including a variant of the Generalized Neighborhood model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a set of measures provided by the Irvine Phonotactic Calculator \citep{Vaden2009}. The most reliable measure is also the most venerable measure of lexical similarity: Coltheart's $N$ \citep{Coltheart1977}, which is defined as the number of words in some representative sample which can be changed into a target nonce word by a single insertion, deletion, or substitution of a phone. \citet{Greenberg1964} find a correlation between wordlikeness ratings and a variant of this measure which only counts words differing by a single substitution. This measure is plotted against ratings from the three studies in Figure \ref{neighborhood}, where it can be seen that it accounts for much of the variance in ratings.

@@

\citet{Hayes2008a} propose a sophisticated model in which well-formedness is related to a probability distribution over phone sequences, estimated according to the principle of maximum entropy, or ``MaxEnt'' \citep[e.g.,][]{Goldwater2003,Jager2007}. \citeauthor{Hayes2008a} use a complex method to evaluate their model. First, they extract onset sequences from the CMU pronunciation dictionary, and use these to train the model. The model is then used to score the onsets of the \citet{Scholes1966} nonce words. Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting correlation.\footnote{This is contrary to standard practices in natural language process, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter); when this is the case, there is reason to suspect the parameter values will not generalize to new data.
No transformation is used here; this only has an effect on the Pearson $r$ coefficient, since they use a transformation that preserves monotonicity.}

An attempt is made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications are all used. However, \citet{Albright2009a} finds that the maximum entropy model, training and testing only on onsets, does not generalize to the \citet{Albright2003b} data, presumably because of the considerable diversity of rimes in the latter data. Consequently, the model was trained to score whole words, not just onsets, using the subset of the CMU dictionary described above. Following the procedure of \citet{HayesInPress}, dictionary entries were syllabified and the resulting syllables were parsed into onset, nucleus, and coda, and a novel feature [$\pm$\textsc{Coda}] is added to contrast onset and coda consonants. Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}. The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to large variety of nonce words, though many of these words appear to have a low rating. This suggests that the model is not robust enough to score whole words.

@@

\begin{table}
\centering
\begin{tabular}{l rrr c rrr c rrr c rrr}
\toprule
              &         \multicolumn{3}{c}{Pearson $r$} &&     \multicolumn{3}{c}{Spearman $\rho$} && \multicolumn{3}{c}{G-K $\gamma$} && \multicolumn{3}{c}{Kendall $\tau_b$} \\
\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12} \cmidrule{14-16}
              &           A &         AH &            S &&           A &       AH    &           S &&           A &          AH &           S &&           A &          AH & S \\
\midrule
Gross status  & \uline{.73} &       {.60} &       {.80} && \uline{.82} &       {.66} &       {.80} && \uline{.87} & \uline{.93} & \uline{.91} &&       {.67} &       {.47} &       {.62} \\
Density ($N$) &       {.67} & \uline{.79} & \uline{.86} &&       {.61} & \uline{.74} & \uline{.82} &&       {.49} &       {.57} &       {.74} &&       {.45} & \uline{.56} & \uline{.67} \\
Bigram score  &       {.46} &       {.75} &       {.74} &&       {.34} &       {.70} &       {.79} &&       {.25} &       {.50} &       {.63} &&       {.25} &       {.50} &       {.61} \\
MaxEnt score  &       {.70} &       {.21} &       {.53} &&       {.66} &       {.39} &       {.58} &&       {.85} &       {.61} &       {.56} && \uline{.68} &       {.16} &       {.48} \\ 
\bottomrule
\end{tabular}
\caption{Correlations between item-averaged wordlikeness ratings and model scores; the largest coefficient for each statistic/data set pair is underlined.}
\label{scores}
\end{table}

@@

A serious limitation of this evaluation is the primitive nature of the gross phonotactic status baseline. It does not allow for any way to state constraints on onset-nucleus sequences, which have been proposed for some languages (e.g., \citealt{Kirby2007}),\footnote{\citet{Kessler1997}, however, argue there are no clear restrictions on English onset-nuclei pairs.} or constraints spanning whole syllables, which have been proposed for English \citep[e.g.,][]{Berkley1994b,Berkley1994a,Coetzee2008b,Fudge1969}. Furthermore, the gross phonotactic baseline does not have any mechanism for generalizing the wellformedness of [ɛsp] rimes from \emph{clasp}, \emph{lisp}, and other rimes consisting of a lax vowel followed by [sp] found in English. \citet{Borowsky1989}, for instance, proposes a theory of possible rimes in English which makes the correct prediction regarding [ɛsp]. This is not embedded in an acquisitional model, but models of syllable type acquisition have been proposed \citep[e.g.,][]{Levelt2000}.

The gross phonotactic baseline could also be extended so as to include additional levels of wellformedness. While the bigram and MaxEnt models do not appear to be able to reliably distinguish intermediate levels of well-formedness, it might be desirable to encode the intuition that, for example, [ʒlɪk] \emph{zhlick}, is more English-like than [bnɪk], though both have unattested onsets. It is also possible to imagine that phonotactic violations would have a cumulative effect on well-formedness; for instance, a nonce word with an unattested onset and an unattested rime, like [tsɪlb], might be less English-like than either [tsɪb] or [sɪlb]. This prediction is made by the bigram and MaxEnt models, among others \citep[e.g.,][]{Albright2008,Anttila2008a}, but could easily be incorporated into a simple baseline by counting the number of violations. However, there is not yet any convincing evidence for cumulativity effects in wordlikeness tasks.

>>>>

Scientific observations do not necessarily arrive at the appropriate granularity for analysis.
Since the granularity of measurement and analysis is determined by prior hypotheses about the nature of the system under study, disputes are inevitable as scientific paradigms evolve. 
This study argues that there are both theoretical and empirical reasons to doubt the implicit hypothesis linking scalar wordlikeness judgements and a gradient model of phonotactics. First, intermediate judgements are characteristic of all gradient rating tasks, and therefore these judgements are irrelevant to the question of whether well-formedness is categorical or gradient.
Secondly, simple baselines better account for gradient well-formedness judgements than current computational models of phonotactic knowledge, suggesting that the gradience observed in these tasks do not derive from grammatical mechanisms.

@@

\section{Aspects of the theory of gradient grammaticality}

@@

\begin{figure}
\centering
\includegraphics{agg.pdf}
\caption{Subjects freely use intermediate ratings when asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively (\citealp{Armstrong1983}; ``1'' indicates ``most representative'')}
\label{agg}
\end{figure}

@@

Each step of this procedure is dubious, however.
First, speakers have difficulty perceiving \citep{Berent2007a,Brown1956,Dupoux1999,Kabak2007a} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
Secondly, the computation of a scalar value serves no further purpose than to provide for gradient well-formedness judgements; even if this is the result of probabilistic parsing \citep[e.g.,][]{Coleman1997}, 
it need not be computed overtly.\footnote{
    An objection might be made here on evolutionary grounds. 
    It is still quite mysterious why the human language endowment includes constraints on pronominal binding, for instance, but these constraints are implicated in everyday language use.
    Far more bizarre is the suggestion that the linguistic endowment includes mechanisms only used in certain experimental tasks.}
Next, speakers must be able to consciously access and report the magnitude of this value (it must be \emph{cognitively penetrable} in the sense of \citealt{Pylyshyn1984}), an ability which is limited in many other domains.
Finally, \citet{Sprouse2011} argues that speakers do not (or cannot) scale well-formedness values to satisfy the assumptions of magnitude estimation, a type of gradient rating task.
%\citep{Bard1996}

It is informative to compare this baroque architecture to a model of what is necessary to make a binary well-formedness judgement. When presented with a linguistic item in a judgement task, the grammar attempts to assign a parse. Speakers then access whether or not parsing was successful. 
There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}. 
As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures may not be stored in linguistic memory. 
%This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
The fact that requests for repetition and clarification are ubiquituous in spontaneous speech illustrates further that speakers are frequently aware when parsing has failed.
Consequently, a great deal of evidence is needed to reject this simple model in favor of the gradient grammaticality architecture.

@@

It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared. 
\citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables. 
Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words.
Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, many of which are exceptionless, into probabilities. 
While this is consistent with their claim, that ``the ability to model gradient intuitions [are] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.

@@

\citet{Scholes1966} conducts several wordlikeness tasks with 7th-grade students (approximately 12--13 years of age). The data used here is his experiment 5, in which 33 speakers provide a ``yes'' or ``no'' as to whether each of the 63 items, presented aurally, are ``likely to be usable as a word of English''. Like the study by \citet{Albright2007}, the focus is on onset well-formedness and there is minimal diversity in rime type. Two items, [klʌŋ] \emph{clung} and [bɹʌŋ] \emph{brung} (a dialectal past participle of \emph{bring}), are excluded here as actual words of English. \citet{Albright2009a} and \citet{Hayes2008a} also use this data for the purposes of model evaluation; following \citet{Frisch2000}, they use the proportion of ``yes'' responses for each item so as to derive a continuous measure of well-formedness.

@@

Models are evaluated by comparing their scores to the average rating of each word using four correlation statistics. Each of these range between $[-1, 1]$, where $1$ indicates a perfect positive correlation and $-1$ denotes a perfect negative correlation. \citet{Hayes2008a} evaluate their model using the Pearson (``product-moment'') $r$, a parametric correlation measure.  \citet{Stevens1946} argues that statistics of this type are inappropriate for analysis of Likert scale data, like those used by \citet{Albright2007} and \citet{Albright2003b}.
The reason is that the Pearson $r$ makes a \emph{linearity assumption}. That is, it assumes that nonce words rated ``1'' and  ``3'', for instance, are just as different as are those rated ``4'' and ``6''. A weaker assumption, more appropriate for Likert scale data, is the \emph{monotonicity assumption}: that ``1'' is less English-like than ``3'', which is less English-like than ``4'', and so on. However, it also has been claimed that $r$ is particularly robust to violations of the linearity assumption \citep[e.g.,][]{Havlicek1976}. Pearson $r$ is reported here, but no stance is taken on its appropriateness for this data.
%\citet{Jamieson2004}
\citeauthor{Hayes2008a} also report Spearman $\rho$; this statistic requires only the weaker assumption of monotonicity, but it is difficult to give a simple interpretation to the coefficient.

Much easier to interpret are two other non-parametric statistics, the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ \citep{Noether1981}, as follows. These statistics are computed by comparing every model score/wordlikeness rating pair to every other: a comparison is counted as \emph{concordant} if the greater of the two model scores is the one associated with the greater of the two wordlikeness ratings (that is, the model ranks these two nonce words in accordance with speakers' ratings), and as \emph{discordant} otherwise. These two statistics differ only in the treatment of ``ties'', pairs where either the model score or wordlikeness rating are identical. For $\gamma$, ties are ignored, and the coefficient is

@@

\noindent
where $c$ and $d$ represent the number of concordant and discordant pairs, respectively. The $\tau_b$ statistic uses a similar formula, but also incorporates a penalty for ties in model score which are not also paired with ties in wordlikeness ratings, or vice versa. \citet{Albright2009a} uses a variant of this statistic to evaluate wordlikeness models.

@@

Operationalizing ``phonotactic violation'' is somewhat more difficult. The simplest possible mechanism is chosen here: an onset or rime is identified as well-formed if it occurs with non-zero frequency in a representative sample, and is identified as ill-formed otherwise. The sample is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms, the latter thought to be particularly strongly correlated with behavioral measures \citep{Brysbaert2009}. These pronunciations are then syllabified, and individual syllables parsed into onset and rime, according to a process described in detail in Appendix \ref{syllabification}. This is not to imply that all unattested onsets or rimes should be regarded as ill-formed, or that all onsets or rimes with non-zero frequency in this data are well-formed. For instance, \citet{Albright2009a} judges [dɹɛsp] \emph{dresp} to be phonotactically well-formed, despite the total lack of [ɛsp] rimes in English. Similar observations have been made concerning English onsets \citep[e.g.,][]{Cairns1972,Moreton2002}. In the other direction, there are precedents for labeling certain attested words as phonotactically ``peripheral'' (e.g., \citealt{Myers1987}, \citealt{Borowsky1989}), as lexical exceptions to language-specific prosodic constraints, without labeling these words as absolutely ungrammatical. Neither of these dissociations between attestation and well-formedness are implemented here, however.

@@

A second baseline is provided by measures of similarity to existing English words, which has long been applied to model wordlikeness judgements (e.g., \citealt{Bailey2001}, \citealt{Greenberg1964}, \citealt{Kirby2007}, \citealt{Ohala1986b}, \citealt{Shademan2006,Shademan2007}, \citealt{Vitevitch1998,Vitevitch1999a}). \citeauthor{LSLT} (\citeyear{LSLT}: 151, fn.~27) suggests that grammaticality judgements in general might be influenced by similarity to existing grammatical structures, and \citet[417f.]{SPE} outline a similarity-based wordlikeness model. More recently, it has been observed (e.g., \citealt[51]{Coleman1997}, \citealt{Hay2004a}) that nonce words which flagrantly violate English sonority restrictions but which bear common affixes (e.g., *\emph{mrupation}) are rated highly English-like.

A wide variety of lexical similarity measures were considered, including a variant of the Generalized Neighborhood model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a set of measures provided by the Irvine Phonotactic Calculator \citep{Vaden2009}. The most reliable measure is also the most venerable measure of lexical similarity: Coltheart's $N$ \citep{Coltheart1977}, which is defined as the number of words in some representative sample which can be changed into a target nonce word by a single insertion, deletion, or substitution of a phone. \citet{Greenberg1964} find a correlation between wordlikeness ratings and a variant of this measure which only counts words differing by a single substitution. This measure is plotted against ratings from the three studies in Figure \ref{neighborhood}, where it can be seen that it accounts for much of the variance in ratings.

@@

\citet{Hayes2008a} propose a sophisticated model in which well-formedness is related to a probability distribution over phone sequences, estimated according to the principle of maximum entropy, or ``MaxEnt'' \citep[e.g.,][]{Goldwater2003,Jager2007}. \citeauthor{Hayes2008a} use a complex method to evaluate their model. First, they extract onset sequences from the CMU pronunciation dictionary, and use these to train the model. The model is then used to score the onsets of the \citet{Scholes1966} nonce words. Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting correlation.\footnote{This is contrary to standard practices in natural language processing, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter); when this is the case, there is reason to suspect the parameter values will not generalize to new data.
No transformation is used here; this only has an effect on the Pearson $r$ coefficient, since they use a transformation that preserves monotonicity.}

An attempt is made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications are all used. However, \citet{Albright2009a} finds that the maximum entropy model, training and testing only on onsets, does not generalize to the \citet{Albright2003b} data, presumably because of the considerable diversity of rimes in the latter data. Consequently, the model was trained to score whole words, not just onsets, using the subset of the CMU dictionary described above. Following the procedure of \citet{HayesInPress}, dictionary entries were syllabified and the resulting syllables were parsed into onset, nucleus, and coda, and a novel feature [$\pm$\textsc{Coda}] is added to contrast onset and coda consonants. Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}. The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to a large variety of nonce words, though many of these words appear to have a low rating. This suggests that the model is not robust enough to score whole words.

@@

\begin{table}
\centering
\begin{tabular}{l rrr c rrr c rrr c rrr}
\toprule
              &         \multicolumn{3}{c}{Pearson $r$} &&     \multicolumn{3}{c}{Spearman $\rho$} && \multicolumn{3}{c}{G-K $\gamma$} && \multicolumn{3}{c}{Kendall $\tau_b$} \\
\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12} \cmidrule{14-16}
              &           A &         AH &            S &&           A &       AH    &           S &&           A &          AH &           S &&           A &          AH & S \\
\midrule
Gross status & \uline{.73} &       {.60} &       {.80} && \uline{.82} &       {.66} &       {.80} && \uline{.87} & \uline{.93} & \uline{.91} &&       {.67} &       {.47} &       {.62} \\
Density ($N$) &       {.67} & \uline{.79} & \uline{.86} &&       {.61} & \uline{.74} & \uline{.82} &&       {.49} &       {.57} &       {.74} &&       {.45} & \uline{.56} & \uline{.67} \\
Bigram $p$ &       {.46} &       {.75} &       {.74} &&       {.34} &       {.70} &       {.79} &&       {.25} &       {.50} &       {.63} &&       {.25} &       {.50} &       {.61} \\
MaxEnt $p$ &       {.70} &       {.21} &       {.53} &&       {.66} &       {.39} &       {.58} &&       {.85} &       {.61} &       {.56} && \uline{.68} &       {.16} &       {.48} \\ 
\bottomrule
\end{tabular}
\caption{Correlations between item-averaged wordlikeness ratings and model scores; the largest coefficient for each statistic/data set pair is underlined.}
\label{scores}
\end{table}

@@

A serious limitation of this evaluation is the primitive nature of the gross phonotactic status baseline. It does not allow for any way to state constraints on onset-nucleus sequences, which have been proposed for some languages (e.g., \citealt{Kirby2007}),\footnote{\citet{Kessler1997}, however, argue there are no clear restrictions on English onset-nuclei pairs.} or constraints spanning whole syllables, which have been proposed for English \citep[e.g.,][]{Berkley1994b,Berkley1994a,Coetzee2008b,Fudge1969}. Furthermore, the gross phonotactic baseline does not have any mechanism for generalizing the wellformedness of [ɛsp] rimes from \emph{clasp}, \emph{lisp}, and other rimes consisting of a lax vowel followed by [sp] found in English. \citet{Borowsky1989}, for instance, proposes a theory of possible rimes in English which makes the correct prediction regarding [ɛsp]. This is not embedded in an acquisitional model, but models of syllable type acquisition have been proposed \citep[e.g.,][]{Fikkert1994,Levelt2000,Pan2003,Pan2004}.

The gross phonotactic baseline could also be extended so as to include additional levels of wellformedness. While the bigram and MaxEnt models do not appear to be able to reliably distinguish intermediate levels of well-formedness, it might be desirable to encode the intuition that, for example, [ʒlɪk] \emph{zhlick}, is more English-like than [bnɪk], though both have unattested onsets. It is also possible to imagine that phonotactic violations would have a cumulative effect on well-formedness; for instance, a nonce word with an unattested onset and an unattested rime, like [tsɪlm], might be less English-like than either [tsɪl] or [sɪlm]. This prediction is made by the bigram and MaxEnt models, among others \citep[e.g.,][]{Albright2008,Anttila1997}, but could easily be incorporated into a simple baseline by counting the number of violations. However, there is not yet any convincing evidence for cumulativity effects in wordlikeness tasks.
